---
title: "Project"
author: "Mark Cunningham, Ethan Saloom, Branson Stickney"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: show
    csl: biomed-central.csl
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    number_sections: yes
    theme: journal
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: 4
  pdf_document:
    df_print: kable
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    highlight: tango
    toc: yes
    toc_depth: 4
---

# Data Preprocessing

## Library Imports

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
library(dplyr)
library(plyr)
library(ggplot2)
library(tidyr)
library(ggpubr)
library(recommenderlab)
library(data.table)
library(reshape2)
library(Matrix)
library(factoextra)
library(knitr)
library(kableExtra)
library(stringr)
library(proxy)
```
## Dataframe Imports

```{r}
anime_df <- read.csv("anime.csv")
ratings_df <- read.csv("rating_complete.csv")
```

## Filtering out explicit content and Subsetting Irrelevant Columns

```{r}
# Filtering out explicit anime from anime_df
# Filtering out anime with no English names, or less than 1000 complete watches
anime_df <- anime_df %>% filter(!(grepl("Hentai",Genres)|Score=="Unknown"|English.name=="Unknown") & anime_df$Completed > 1000) 

# Updating ratings dataframe with anime datframe
ratings_df <- ratings_df[(ratings_df$anime_id %in% anime_df$MAL_ID),]

anime_df <- subset(anime_df, select=-c(Japanese.name,Producers,Licensors,Favorites,Watching,On.Hold,Plan.to.Watch,Score.10,Score.9,Score.8,Score.7,Score.6,Score.5,Score.4,Score.3,Score.2,Score.1))
```

```{r}
#Old code:
#hentai_df <- anime_df %>% filter(grepl("Hentai",Genres)|Score=="Unknown")
#anime_df <- anime_df %>% filter(Score!="Unknown") %>% filter(!grepl("Hentai",Genres))
#ratings_df<-ratings_df[!(ratings_df$anime_id %in% hentai_df$MAL_ID),]  


#More old code:
#Filter only anime with greater than 200 ratings and users with 200 ratings
#ratings_df <- ratings_df %>% 
#  group_by(anime_id) %>%
#  filter(n()>200) %>%
#  ungroup() %>%
#  group_by(user_id) %>%
#  filter(n()>400)  %>%
#  ungroup()

#Group by user_id, sample 5000 users, ungroup
#ratings_df <- ratings_df %>% group_by(user_id,)
#user_sample <- sample(nrow(ratings_df), 5000)
#ratings_df_sample <- ratings_df[user_sample,]
#ratings_df_sample %>% ungroup()
```

## Previewing the Data

```{r}
kable(head(anime_df, n = 10L)) %>%
  column_spec(7,width="3in") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"),full_width = F) %>%
  row_spec(0,bold = T, color = "black", background = "#fc5e5e") %>%
  row_spec(1:10,bold = T, color = "black") %>%
  column_spec(1:17,color="black",width="18em") 

kable(head(ratings_df, n = 10L)) %>%
  kable_styling(bootstrap_options = c("striped", "condensed")) %>%
  row_spec(0,bold = T, color = "white", background = "#fc5e5e") %>%
  scroll_box(width = "100%", height = "200px")

anime_df %>%
  ggplot(aes(x=Type)) +
  geom_bar(stat = "count", fill = "#fc5e5e" ) +
  ggtitle("Anime Type Distribution")
```

## Sampling

```{r}
#Unlist data so it can be transformed into realRatingMatrix
ratings_df$user_id <- as.numeric(unlist(ratings_df$user_id))
ratings_df$anime_id <- as.numeric(unlist(ratings_df$anime_id))
ratings_df$rating <- as.numeric(unlist(ratings_df$rating))
ratings_df<-as.data.frame(ratings_df)

set.seed(1234)
#Convert rating matrix into a recommenderlab sparse matrix
ratingmat <- as(ratings_df, "realRatingMatrix")
ratingmat
#Filter for anime that has 1000 or more ratings, and users who have greater than 100 anime reviews
ratingmat <- ratingmat[rowCounts(ratingmat) > 100, colCounts(ratingmat) > 1000]
ratingmat
object.size(ratingmat)
```


```{r}
set.seed(1234)
#Sample of rating matrix
ratingmat_sample <- sample(ratingmat, 5000)
ratingmat_sample
object.size(ratingmat_sample)
```

## Preview Rating Distribution

```{r}
#Histograms
hist(getRatings(ratingmat_sample), breaks=100)
hist(getRatings(normalize(ratingmat_sample,method="Z-score")),breaks=100)
hist(rowCounts(ratingmat_sample),breaks=50,xlim=range(0:1000))
```


## Top Ten Most Viewed Anime in Sample

```{r}
views_per_anime <- colCounts(ratingmat_sample)
table_views <- data.frame(anime_names = names(views_per_anime), views = views_per_anime)
names(table_views)[names(table_views) == "anime_names"] <- "anime_id"
ratingmat_sample_df <- as(ratingmat_sample, "data.frame")
anime_df_sample <- anime_df[(anime_df$MAL_ID %in% ratingmat_sample_df$item),]
anime_df_sample = dplyr::rename(anime_df_sample, anime_id = MAL_ID)

table_views <- merge(table_views, anime_df_sample, by="anime_id")
table_views <- table_views[order(table_views$views, decreasing = TRUE), ]

ggplot(table_views[1:10, ], aes(x = reorder(Name, -views), y = views)) + 
  xlab("") + ylab("Number of Participants") + xlab("Anime Name") +
  geom_bar(stat = "identity") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))+ 
  ggtitle("Top 10 Viewed Anime in Sample")

```

## Similarity

```{r}
image(ratingmat_sample[1:100, 1:100], main = "First 100 users and anime items: Top Anime (Non-Normalized)")

ratingmat_norm <- normalize(ratingmat_sample)
avg <- round(rowMeans(ratingmat_norm), 5)

image(ratingmat_norm[1:100, 1:100], main = "First 100 users and anime items: Top Anime (Normalized)")

similarity_users <- similarity(ratingmat_norm[1:100,], 
                               method = "cosine", 
                               which = "users")
image(as.matrix(similarity_users), main = "User Similarity")

similarity_anime <- similarity(ratingmat_norm[,1:100], 
                               method = "cosine", 
                               which = "items")
image(as.matrix(similarity_anime), main = "Anime Similarity")
```



# Creating the Algorithms

## Splitting the Data

```{r}
min(rowCounts(ratingmat_sample))
```

```{r}
anime_eval <- evaluationScheme(data = ratingmat_sample, method = "split", 
                                     train = 0.8, given = min(rowCounts(ratingmat_sample)), goodRating = 7) 
anime_eval
```

## Which models are available in recommenderlab?

```{r}
recommender_models <- recommenderRegistry$get_entries(dataType ="realRatingMatrix")
recommender_models$IBCF_realRatingMatrix$parameters
recommender_models$UBCF_realRatingMatrix$parameters
```

## Item-Based Collaborative Filtering

```{r}
ibcf_model <- Recommender(getData(anime_eval,"train"), 
                          method = "IBCF")
ibcf_model
```

## User-Based Collaborative Filtering

```{r}
ubcf_model <- Recommender(getData(anime_eval,"train"), 
                          method = "UBCF")
ubcf_model
```

```{r}
ibcf_predict <- predict(ibcf_model, getData(anime_eval, "known"), type="ratings")
ubcf_predict <- predict(ubcf_model, getData(anime_eval, "known"), type="ratings")

error <- rbind(IBCF = calcPredictionAccuracy(ibcf_predict, getData(anime_eval, "unknown")),
               UBCF = calcPredictionAccuracy(ubcf_predict, getData(anime_eval, "unknown")))
error

#image(getModel(ibcf_model)$sim,
      #main = "Heatmap of the first rows and columns")
#image(getModel(ubcf_model)$sim,
#     main = "Heatmap of the first rows and columns")
```

## What all can we do with the models?

```{r}
names(getModel(ubcf_model))
names(getModel(ibcf_model))
```

## Recommendations for Users

RecommenderLab creates predictions using predict() and creates an object of class TopNList.

### UBCF

```{r}
ubcf_model_test <- Recommender(ratingmat_sample,method="UBCF")
recom<-predict(ubcf_model_test, ratingmat_sample[1],n=5)
#class(recom)

getAnimeNames <- function(df) {
  out = as.data.frame(matrix(ncol=1,nrow=5))
  names(out)[1] <- 'Anime Name'
  for(i in 1:nrow(df)) {
      value <- df[[i,1]]
      row <- which(anime_df[["MAL_ID"]]==value)
      out[i,1] <- anime_df[row,5]
  }
  return(out)
}
recommend_example_ubcf <- getAnimeNames(as.data.frame(as(recom,"list")))
recommend_example_ubcf 
getList(bestN(recom, 3))[[1]]
getList(recom)
```

### IBCF

```{r}
recom<-predict(ibcf_model, ratingmat[1001],n=5)
recommend_example_ibcf <- getAnimeNames(as.data.frame(as(recom,"list")))
recommend_example_ibcf 

getList(recom)
```


```{r}
recom_list <- as(recom,"list")
getAnimeNames(as.data.frame(as(recom,"list")))
getList(recom)
```


```{r}
#anime_1_id <- which(anime_df_sample[,5]== anime_1)
  #anime_2_id <- which(anime_df_sample[,5]== anime_2)
  #anime_3_id <- which(anime_df_sample[,5]== anime_3)
 # anime_4_id <- which(anime_df_sample[,5]== anime_4)
  #anime_5_id <- which(anime_df_sample[,5]== anime_5)
  user <- c(9,9,9,9,9)
  item <- c(16498, 1535, 39587, 47,30276)
  rating <- c(9,8,9,8,7)
  user_ratings <- data.frame(user,item, rating)
  #row_num <- which(movies2[,2] == input)
  user_ratings$item <- as.character(unlist(user_ratings$item))
  user_ratings$rating <- as.numeric(unlist(user_ratings$rating))
  user_ratings<-as.data.frame(user_ratings)
  ratingmat_sample_df <- as(ratingmat_sample, "data.frame")
  ratingmat_sample_df$user <- as.numeric(ratingmat_sample_df$user)
  ratingmat_sample_df$item <- as.character(ratingmat_sample_df$item)
  ratingmat_sample_df<-ratingmat_sample_df[order(as.numeric(row.names(ratingmat_sample_df))),]
  user_rating_mat5 <- rbind(user_ratings,ratingmat_sample_df)
  user_rating_mat5$user <- as.numeric(user_rating_mat5$user)
  user_rating_mat5$item <- as.character(user_rating_mat5$item)
  user_rating_mat6<- as(user_rating_mat5, "realRatingMatrix")
```


```{r}
  user_rating_mat <- dcast(ratingmat_sample_df, user~item, value.var = "rating", na.rm=FALSE)
  user_rating_mat2 <- user_rating_mat[,-1]
  tmp = matrix(NA,1,3140)
  colnames(tmp) <- colnames(user_rating_mat2)
  tmp[1,"16498"] <- 7
  tmp[1,"1535"] <- 7
  tmp[1,"39587"] <- 7
  tmp[1,"47"] <- 7
  tmp[1,"30276"] <- 7
  user_rating_mat3 <- rbind(tmp,user_rating_mat2)
  user_rating_mat3 <- proxy::as.matrix(user_rating_mat3)
```


```{r}
#Convert rating matrix into a sparse matrix
  # user_rating_mat3<- as(user_rating_mat3, "realRatingMatrix")
  # temp1 <- dcast(ratingmat_sample_df, user~item, value.var = "rating", na.rm=FALSE)
  # tmp = matrix(NA,1,3140)
  # temp2 <- temp1[,-1]
  # colnames(tmp)=colnames(temp2)
  # tmp[1, user_ratings$item] = user_ratings$rating
  # user_rating_mat <- rbind(tmp,temp2)
  # user_rating_mat <- as.matrix(user_rating_mat)
  # user_rating_mat <- as(user_rating_mat,"realRatingMatrix")
  
recommender_model <- Recommender(user_rating_mat6, method = "UBCF")
recom <- predict(recommender_model, user_rating_mat6[1], n=5)
recom_list <- as(recom, "list")
getAnimeNames(as.data.frame(as(recom,"list")))
```


```{r}
as(user_rating_mat3,data.frame)
user_rating_mat3[1]
```

